Draft-assinment1

### 2017-02-20 Draft-xuezhou-sjb.Rmd
update

#Table of Contents
* 1.[Introduction](#Introduction)
* 2.[Goal](#Goal)
* 3.[Dataset Preparation](#dataprep)
* 4.[Objective](#objective)
* 5.[Summary of variables](#variablesummary)

 * 5.1 [Individual variable summary](#5.1)
  * 5.1.1 [School](#5.1.1)
    * 5.1.1.1 [State](#5.1.1.1)
    * 5.1.1.2 [Predominant Degree](#5.1.1.2)
    * 5.1.1.3 [Control of Institution](#5.1.1.3)
  * 5.1.2 Student Body
    * 5.1.2.1 [Percentage first-generation students](#5.1.2.1)
    * 5.1.2.2 [Number of applications](#5.1.2.2)
  * 5.1.3 Cost 
    * 5.1.3.1 [Average net price for Title IV institutions(both public and private institutions)](#5.1.3.1)
    * 5.1.3.2 [Percentage of students who are financially independent](#5.1.3.2)
  * 5.1.4 Financial Aid
    * 5.1.4.1 [Percentage of undergraduates who receive a Pell Grant](#5.1.4.1)
    * 5.1.4.2 [Percent of all federal undergraduate students receiving a federal student loan](#5.1.4.2l)
    * 5.1.4.3 [The median debt for students who have completed](#5.1.4.3)
    * 5.1.4.4 [The median debt for students who have not completed](#5.1.4.4)
  * 5.1.5 Earning
    * 5.1.5.1 [Median family income in real 2015 dollars](#5.1.5.1)
    * 5.1.5.2 [Earning after graduation](#5.1.5.2)
  * 5.1.6 Repayment
    * 5.1.6.1 [3-year Repayment rate](#5.1.6.1)
    * 5.1.6.2 [3-year Default rate](#5.1.6.2)
* 6.Relationship between variables
 * 6.1 [The relationship between NPT4_COMBINE and State](#6.1)
 * 6.2 [The relationship between Debt and Cost](#6.2)
 * 6.3 [The relationship between Percentage of Pell Grant and college type(CONTROL)](#6.3)
 * 6.4 [The relationship between Percentage first-generation students and States](#6.4)
 * 6.5 [The relationship between family income and percentage of Pell Grant](#6.5)
 * 6.6 [The relationship between DEP_STAT_PCT_IND and Control](#6.6)
 * 6.7 [The relationship between PCTLOAN and Control](#6.7)
 * 6.8 [cost and median debts with earning](#6.8)
 * 6.9 [default with earning](#6.9)
 * 6.10 [degree type vs. earning](#6.10)
 * 6.11 [predominant degree, control  and earning after 6 year of entry](#6.11)
 * 6.12 [predominant degree + control (count)](#6.12)
 * 6.13 [control + earning](#6.13)
* 7. [Findings and Conclusion](#7)

@@

#2. Goal<a id="Goal"></a>
#3. Dataset Preparation<a id="dataprep"></a>
Our analysis focuses on the data of year 2012 -2013 of 23 variables. We chose the median income 6 years after graduation as our predictor variable, and eliminated many fields with bunch of missing values as well as some fields not pertainig enougn to post-graduation income. We then fosued primarily on a set of 23 variables, including gender, age, ethnic, acedemic costs, family income, completetion rates and replayment rate.

@@

Considering there is some special characters in the ID coloumn, we first rename the first column's name to be selectable. 
```{R}
colnames(data)[1] = 'UNITID'
```

@@

variable = select(data, UNITID, INSTNM, STABBR, PREDDEG, LOCALE2, MENONLY, WOMENONLY, CONTROL, ADM_RATE, SATMTMID, SATVRMID, SATWRMID, SAT_AVG,
                  NPT4_PUB,
                  NPT4_PRIV,
                  UG,
                  UG_NRA,
                  UG_UNKN,
                  UG_WHITENH,
                  UG_BLACKNH,
                  UG_API,
                  UG_AIANOLD,
                  UG_HISPOLD,
                  DEP_STAT_PCT_IND,
                  PAR_ED_PCT_1STGEN,
                  MD_FAMINC,
                  MEDIAN_HH_INC,
                  POVERTY_RATE,
                  PCTPELL,
                  PCTFLOAN,
                  GRAD_DEBT_MDN,
                  WDRAW_DEBT_MDN,
                  APPL_SCH_PCT_GE2,
                  C150_4,
                  MD_EARN_WNE_P6,
                  MD_EARN_WNE_P10,
                  RPY_3YR_RT_SUPP,
                  CDR3)

@@

#need to be commented out 
class(as.data.frame(apply(variable[,-c(2,3,4,8)],2,as.numeric)))

#get rid of factor variables: INSTNM, STABBR, PREDDEG, CONTROL and change the data types to numeric 
variable[,-c(2,3,4,8)] %>%
    apply(2, as.numeric) %>%
    as.data.frame() %>%
    {.} -> variable_num

#change the categorical level of the categorical variables:PREDDEG, CONTROL
variable$preF=factor(variable$PREDDEG,levels=c(0,1,2,3,4),labels = c("NotClassified","Certificate-degree","Associate's-degree","Bachelor's-degree","Graduate-degree"))

variable$ctrF=factor(variable$CONTROL,levels = c(1,2,3),labels = c("Public","Private Nonprofit","Private For-profit"))

#combine the three datafram to generate a new dataset. 
data_new <- cbind(variable_num, variable$preF, variable$ctrF,variable$INSTNM, variable$STABBR)

#rename the categorical varaibles to approriate readable names
colnames(data_new)[35] = 'PREDDEG_factor'
colnames(data_new)[36] = 'Control_factor'
colnames(data_new)[37] = 'INSTNM_factor'
colnames(data_new)[38] = 'STABBR_factor'

#need to be commented out
mean(is.na(data_new$NPT4_PUB))
```

@@

for (i in 1:length(data_new$NPT4_PRIV)) {
   if (is.na(data_new$NPT4_PRIV[i])) {
       data_new$NPT4_PRIV[i] = data_new$NPT4_PUB[i]
   }
}
mean(is.na(data_new$NPT4_PRIV))
data_new = data_new[,-10] 
colnames(data_new)[10] <- "NPT4_COMBINE"

mean(is.na(data_new$NPT4_COMBINE))
```

In order to choose the variables that have enough valid data related to the topic, we filtered and considered variables with percent missing value less than 25%. The new data set was named as ???data_clean??? for further analysis. 
```{r, message=FALSE,warning=FALSE}

data_clean = data_new[,colMeans(is.na(data_new)) < 0.25]

#rearrange the coloumn names
data_clean = data_clean[c(1,17,18,15,16,2:14)]

write.csv(data_clean,"data_clean.csv")
```
The data_clean will be the cleaned dataset we used in the further variable summaries and comparisons

@@

The table below lists the variables that will be used for the further analysis: 12 of them are continuous variables while 3 of them are categorical variables. Our data set now contains 120 rows and 15 columns. 120 rows consist of 20 years??? of data for each of the chosen six countries. The 15 columns consist of country name, country code, year, and the twelve indicators.

| Variable Name    | Data Type |  Unit  |  Description                                                          |    Category in Documentation       |
| --------------   | ----------|--------|-----------------------------------------------------------------------| --------------------------------|
| NPT4_COMBINE    | Numeric   | $         |average net price for TITLE IV institutions of both public and private.|  Cost                       |
| DEP_STAT_PCT_IND | Numeric  | $         |Percentage of students who are financially independent                |     Cost                        |
| PAR_ED_PCT_1STGEN| Numeric  | %         |Percentage first-generation students                                  |      Sudent Body                |
| MD_FAMINC        | Numeric  | $         |Median family income in real 2015 dollars                             |       Earning                   |
| PCTPELL          | Numeric  | %         |Percentage of undergraduates who receive a Pell Grant                 |        Financial Aid            |
| PCTFLOAN         | Numeric  | %         |Percent of all federal undergraduate students receiving a federal student loan      |  Financial Aid    |
| GRAD_DEBT_MDN    | Numeric  | $         |The median debt for students who have completed                       |   Financial Aid                 |
| WDRAW_DEBT_MDN   | Numeric  | $         |The median debt for sutdents who have not completed                   |Financial Aid |
| MD_EARN_WNE_P6   | Numeric  | $         |Median income six years after entry                                   |     earning   |
| MD_EARN_WNE_P10  | Numeric  | $         |Median income ten years after entry                                   |     earning  |
| RPY_3YR_RT_SUPP  | Numeric  | %         |Repayment rate 3 years after entering repayment                       |Repayment   |
| CDR3             | Numeric  | %         |Loan default rate 3 years afer entering repayment                     |Repayment        |
| PREDDEG_factor    | Factor  | N/A          |Predominant degree offered                                            |School       |
| Control_factor    | Factor  | N/A          |Control of institution                                                |School     |
| STABBR_factor  | Factor     | N/A       |State                                                                 |School     |
| APPL_SCH_PCT_GE2 | Numeric  | %         |Percentage of students submitting 2 or more than 2 applications       | Student        |
| UNITID | Numeric | N/A      | Unique ID for the institutions | unknown |
| INSTNM | Factor |  N/A      |name for the institutions       | unknown |

we will summarize the data set using dplyr and ggplot funtions in the variables summary section.

#4. Objective<a id="objective"></a>
Over years, Many students see college as an investment to help them earn more and live better lives after graduation. Amount of findints have been conducted to analyze the influence of college quality on students' post graudation earning based on *colleges characterestis, student's demographic information, as well as family charateristics as well as federal finanical support to low-income students.  It has become more and more important to understand the fac- tors that contribute to post-graduation earnings and the ability to repay student loans.

Our objective in this stage is to conduct an exploration analysis on the CollegeScore dataset, investigating the individual variable's summary statistics and the relationship between these variabls, so as to better prepare for further analaysis on the factors affecting students' post college earnings with some advanced data mining techniques.

#5. Summary of variables<a id="variablesummary"></a>
##5.1 Individual variable summary

###5.1.1 School
####5.1.1.1. State
State variable has 59 levels, which include 55 states in the United States, as well as......... From the following frequncy table, we see that California has the most University-equivalent institutions across the nation, followed by Texas and Ner York. It has about 800 universities! A bar chart is plotted below to show the top states with the most universities. Then summary of frequency table also shows that the state having the least university only has one, and 50% of state has more than 132 universities in the state.
```{r, message=FALSE}

state_freq = count(data_clean,"variable$STABBR")[order(-count(data_clean,'variable$STABBR')$freq),]
state_freq[1:15,]
summary(state_freq)
state_freq[1:15,] %>% ggplot(aes(x=factor(variable.STABBR), y=freq)) +
  geom_bar(stat = 'identity')

####5.1.1.2. Predominant Degree
Then we analyze the type of degree universities mainly offer. There are four types of degree taken into accout: Certificate-degree, Associate's degree, Bachelor's degree, and Graduate Degree. About 5% of the population are classified in terms of the predomiant degree they offer.  43.2% offer Certificate degree as their predominant degree, 27.3% predominantly offer Bachelor degree and only 4.23% are classified as Graduate degree predominant.

```{r, message=FALSE}
prop.table(table(data_clean$`variable$preF`)) * 100
data_clean %>% ggplot(aes(x=variable$`preF`, fill = 'blue')) + 
  geom_histogram(stat = "count") +
  xlab("Degree") 
```

####5.1.1.3. Control of Institution
In addition to State and predominant degree, control of institution is another categorical variable in the dataset. The label of Control is to indicate whether or not a institution is institution is privately owned and whether or not it is a for-profit organization. There are three levels of this variable: Public, Private Nonprofit, and Private-profit. Public school is assumed to be non-profit so it was not further divided into two subgroups. Below we show the percentage summary of each category and a bar char is plotted to facilitate understanding. About 50% out of all institution are private for-profit university, which is the *most dominant type*. Public and Private Nonprofit are pretty tie in terms of number of observation in each category.

@@

###5.1.2 Student Body
####5.1.2.1 Percentage first-generation students<a id="5.1.2.1"></a>
This element identifies the share of federally aided students at the institution who report that they are first generaion on the FAFSA (Free Application for Federal Student Aid). First generation students are defined as sudents in which neither parents completed college.

With the followin codes, we extracted the institions with the highest and lowest percentage of first-generation students. Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.
```{R, message=FALSE, warinig = FALSE}

#the institions with the highest and lowest percentage of first-generation students
data_clean %>%
 slice(c(which.min(PAR_ED_PCT_1STGEN),which.max(PAR_ED_PCT_1STGEN))) %>%
  select (c(1,2,3,8))

@@

```
From the output, we can tell that the highest value existed in  La Belle Beauty Academy with the value of 95.25%. The lowest value is 6.87%, which existed in Wheaton College. From the mean value of 48.08%, we can tell that about for federally aided students in most colleges, around half (50.8%) of the students' parents did not complete the college. The standard deviation value of 12.2% indicated a relative high difference of this element among different colleges.

@@

####5.1.2.2 Number of applications<a id="5.1.2.2"></a>
The variable APPL_SCH_PCT_GE2 is to describe the percentage of students who list two and more than two institutions when he apply for financial aid when apply for universities. This information is only available to students who submit their applications for federal aid, and not cover the whole student body. Nevertheless, this variable still can provides some useful insights about how many universities a student usually apply to.

The summry statistics show that median percentage of student submitting two or more than two application is 41.67 across all univerisities. The institution having the least percentage only has 4% of student applying to more than one university, while the school with the highest percent has as high as 90% of student applying to more than one university. The boxplot compare the the percent of student applying to only one university and that of student applying to more than 1. We can clearly see that a larger portion of students only list one instituiton when submitting their application for financial aid.

@@

###5.1.3 Cost<a id="5.1.3"></a>
####5.1.3.1 Average net price for Title IV institutions(both public and private institutions)<a id="5.1.3.1"></a>

With the followin codes, we extracted the IV institutions with the highest and lowest Average net price. Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.

data_clean %>%
 slice(c(which.min(NPT4_COMBINE),which.max(NPT4_COMBINE))) %>%
  select (c(1,2,3,6))

@@

```
From the summary results, we learned that the maximum value of the variable is $87,414, which is for Aerosim Flight Academy located in FL. while the minimum value is about $-19,360 - the value for Berea College located in KY. Considering the elements are collected from Title IV institutions, we reasoned that the negative values may be due to the fact that some intituions's finainal aid supoort can cover the students' all the costs, meanwhile overbalancing the cost amount. There are not much differerences between mean value and median value, which is around $15,000. We can conclude that the average net price costs for most institutuins are falling into the this range. The standard deviation value is 7604.884 which is relatively large, indicating a great spread in the data.

@@

####5.1.3.2 Percentage of students who are financially independent<a id="5.1.3.2"></a>

With the followin codes, we extracted the institions which have the highest and lowest percentage of students who are financially independent. Then we observed this varibles's basic summary statistics: min, max, mean, median and standard deviation.
```{R}

data_clean %>%
 slice(c(which.min(DEP_STAT_PCT_IND),which.max(DEP_STAT_PCT_IND))) %>%
  select (c(1,2,3,7))

@@

```
From the summary results, we learned that the maximum value of the variable is 98.56%(Capella University) while the minimum value is about 0.73%(University of Notre Dame). From the mean value, we can tell that about in most colleges, around 56% of the students are financially independent. The standatd deviation value of 23.3% indicates a relative high different among different colleges, which can also be concluded from the difference between maximum value and minimum value. Below is the code to generate a visual representation in the histgram.

@@

###5.1.4 Financial Aid 
####5.1.4.1 Percentage of undergraduates who receive a Pell Grant<a id="5.1.4.1"></a>

This elemnt, pulled from the IPEDS student Financial Aid Component, shows the share of undergraduate students who received Pell Grants in a given year, which is an important measure of the access an insiitution provides to low-income students.

We are using the code below to summarize this variable and extraced the instition with the highest and lowest value.

@@

data_clean %>%
 slice(c(which.min(PCTPELL),which.max(PCTPELL))) %>%
  select (c(1,2,3,10))

@@

```
From the output, we can tell that the range is between 0 to 100%.	No students in Thunderbird School of Global Management has received the Pell grant while all of the students in American Conservatory Theater have receiced the Pell Grant. Among the colleges in the sample, half colleges have this rate as high as 52.7% (median value of 52.7%), which means more than half of undergraduare students are able to receive the Pell Grant to cover their costs. The standard deviation value is 22.57%,  indicating a relative high difference of this element among different colleges, which can also be noticed from the largest range (100%). Below is the code to generate a visual representation in the histgram.

@@

####5.1.4.2 Percent of all federal undergraduate students receiving a federal student loan<a id="5.1.4.2"></a>

This elemnt, as reported in the IPEDs Student Finaicial Aid component, show the share of undergraduate students who received federal loans in a given year. It can provide important context to figures related to debt, repayment and non-repayment.  
We are using the code below to summarize this variable and extracted the institutions with highest and lowest value.

@@

data_clean %>%
 slice(c(which.min(PCTFLOAN),which.max(PCTFLOAN))) %>%
  select (c(1,2,3,11))

@@

```
From the output, we can tell that the range is between 0 to 100%.	As we can recall from the previous variable(PCTPELL), American Conservatory Theater has both the PCTLOAN and PCTPELL the value of 100%. Gadsden State Community College has no students reveiced federal loans in the given year. Among the colleges in the sample, half colleges have this rate as high as 60.08% , which means around 60% of undergraduare students are able to receive the fedaral student loans. The standard deviation value is 29.26%,  indicating a relative high difference of this element among different colleges, which can also be noticed from the largest range (100%). Below is the code to generate a visual representation in the histgram.

@@

####5.1.4.3 The median debt for students who have completed<a id="5.1.4.3"></a>

This element is the median loan debt acculated at the institution by all student borrowers of federal lonans who have habe completed the education. We are using the code below to summarize this variable. 
```{R, message=FALSE, warinig = FALSE}

data_clean %>%
 slice(c(which.min(GRAD_DEBT_MDN),which.max(GRAD_DEBT_MDN))) %>%
  select (c(1,2,3,12))

@@

```
From the output, we can tell that the highest value is 48,250 - Southwest University of Visual Arts-Tucson, while the lowest value is 1448 - South Florida Institute of Technology. From the median value of 14000, we can tell that among the colleges in the sample, the students who have completed the deducaiton from half of them had the median loan debt of 14000. The standard deviation value is 8020.958,  indicating a relative high difference of this element among different colleges. Below is the code to generate a visual representation in the histgram.

@@

####5.1.4.4 The median debt for students who have not completed<a id="5.1.4.4"></a>
WDRAW_DEBT_MDN describe the median number of debt fot students who ahve not completed their study. In other words, they are still attending the university. The summary statistics show that student in 50% of institution have a median debt of 5500.The median debt for attending students can be as much as 27620 for some school and as low as 600 for other. So we also calculate standard deviation to measure the dispenser of the distribution. The standard deviation is 2415.623 about 95% of institution have a median debt ranging from 1167.754  to 10830.246. The histogram that the distribution of the median roughly follows a standard distribution with a sharp peak and slightly heavy tail.
```{r, message=FALSE}
median_debt_non_null = data_clean$WDRAW_DEBT_MDN[!is.na(data_clean$WDRAW_DEBT_MDN)]
summary(median_debt_non_null)
sd(median_debt_non_null)

@@

###5.1.5 Earning
####5.1.5.1 Median family income in real 2015 dollars<a id="5.1.5.1"></a>
This elemnt is recording the median family income of students, one of the undergraduate student demographics for earnings cohorts.

@@

data_clean %>%
 slice(c(which.min(MD_FAMINC),which.max(MD_FAMINC))) %>%
  select (c(1,2,3,9))

@@

#6. Relationship between variables<a id="6"></a>

##6.1. The relationship between NPT4_COMBINE and State<a id="6.1"></a>

>>>>

#Table of Contents
* 1.[Introduction](#Introduction)
* 2.Goal

College is an investment. It's well-known that a major motivation for students to attend a university is higher post-graduation earning. So will your school make sure you get a return? With collge costs and student-debt levels reaching an all-time high, it's important to know which schools produce the highest earning alumni and what factors can help us predict post-graduate earnings for students in a university.

The goal of our study is to understand what factors are predictive of students' post-graduate earning through individual variable analysis and relationship detection. To this end, we first explore a set of variables that are carefully selected to represent a school's profile, getting to know the distribution of each variable. Then we further analyze how these variables are related, particually the relationship between earning and other variables.

* 3.Dataset Preparation
* 4.Objective
* 5.Summary of variables
 * 5.1 Individual variable summary
  * 5.1.1 School
    * 5.1.1.1 State
    * 5.1.1.2 Predominant Degree
    * 5.1.1.3 Control of Institution
  * 5.1.2 Student Body
    * 5.1.2.1 Percentage first-generation students
    * 5.1.2.2 Number of applications
  * 5.1.3 Cost 
    * 5.1.3.1 Average net price for Title IV institutions(both public and private institutions)
    * 5.1.3.2 Percentage of students who are financially independent
  * 5.1.4 Financial Aid
    * 5.1.4.1 Percentage of undergraduates who receive a Pell Grant
    * 5.1.4.2 Percent of all federal undergraduate students receiving a federal student loan
    * 5.1.4.3 The median debt for students who have completed
    * 5.1.4.4 The median debt for students who have not completed
  * 5.1.5 Earning
    * 5.1.5.1 Median family income in real 2015 dollars
    * 5.1.5.2 Earning after graduation
  * 5.1.6 Repayment
    * 5.1.6.1 3-year Repayment rate
    * 5.1.6.2 3-year Default rate
* 6.Relationship between variables
 * 6.1 NPT4_COMBINE VS State
 * 6.2 Debt VS Cost
 * 6.3 Percentage of Pell Grant VS college type(CONTROL)
 * 6.4 Percentage first-generation students VS Statu
 * 6.5 family income with Pell Grant
 * 6.6 cost and median debts with earning
 * 6.7 default with earning
 * 6.8 degree type vs. earning
 * 6.9 predominant degree, control  and earning after 6 year of entry
 * 6.10 predominant degree + control (count)
 * 6.11 control + earning
* 7. Comparison between variables
* 8. Findings and Conclusion

@@

#2. Goal
#3. Dataset Preparation
Our analysis focuses on the data of year 2012 -2013 of 23 variables. We chose the median income 6 years after graduation as our predictor variable, and eliminated many fields with bunch of missing values as well as some fields not pertainig enougn to post-graduation income. We then fosued primarily on a set of 23 variables, including gender, age, ethnic, acedemic costs, family income, completetion rates and replayment rate.

@@

@@

variable = select(data, STABBR, PREDDEG, LOCALE2, MENONLY, WOMENONLY, CONTROL, ADM_RATE, SATMTMID, SATVRMID, SATWRMID, SAT_AVG,
                  NPT4_PUB,
                  NPT4_PRIV,
                  UG,
                  UG_NRA,
                  UG_UNKN,
                  UG_WHITENH,
                  UG_BLACKNH,
                  UG_API,
                  UG_AIANOLD,
                  UG_HISPOLD,
                  DEP_STAT_PCT_IND,
                  PAR_ED_PCT_1STGEN,
                  MD_FAMINC,
                  MEDIAN_HH_INC,
                  POVERTY_RATE,
                  PCTPELL,
                  PCTFLOAN,
                  GRAD_DEBT_MDN,
                  WDRAW_DEBT_MDN,
                  APPL_SCH_PCT_GE2,
                  C150_4,
                  MD_EARN_WNE_P6,
                  MD_EARN_WNE_P10,
                  RPY_3YR_RT_SUPP,
                  CDR3)

@@

class(as.data.frame(apply(variable[,-c(1,2,6)],2,as.numeric)))

#numeric
variable[,-c(1,2,6)] %>%
    apply(2, as.numeric) %>%
    as.data.frame() %>%
    {.} -> variable_num

#categorical
variable$preF=factor(variable$PREDDEG,levels=c(0,1,2,3,4),labels = c("NotClassified","Certificate-degree","Associate's-degree","Bachelor's-degree","Graduate-degree"))
variable$ctrF=factor(variable$CONTROL,levels = c(1,2,3),labels = c("Public","Private Nonprofit","Private For-profit"))

#combine the three datafram to generate a new dataset. 
data_new <- cbind(variable_num, variable$preF, variable$ctrF,variable$STABBR)
dim(data_new)

mean(is.na(data_new$NPT4_PUB))
```

@@

for (i in 1:length(data_new$NPT4_PRIV)) {
   if (is.na(data_new$NPT4_PRIV[i])) {
       data_new$NPT4_PRIV[i] = data_new$NPT4_PUB[i]
   }
}
mean(is.na(data_new$NPT4_PRIV))
data_new = data_new[,-9] 
colnames(data_new)[9] <- "NPT4_COMBINE"
```

In order to choose the variables that have enough valid data related to the topic, we filtered and considered variables with percent missing value less than 25%. The new data set was named as “data_clean” for further analysis. 
```{r, message=FALSE,warning=FALSE}

data_clean = data_new[,colMeans(is.na(data_new)) < 0.25]
write.csv(data_clean,"data_clean.csv")
```

@@

The table below lists the variables that will be used for the further analysis: 12 of them are continuous variables while 3 of them are categorical variables. Our data set now contains 120 rows and 15 columns. 120 rows consist of 20 years’ of data for each of the chosen six countries. The 15 columns consist of country name, country code, year, and the twelve indicators.

| Variable Name    | Data Type         | Description                                                          |    Category in Documentation       |
| --------------  | ------------------ | ---------------------------------------------------------------------| --------------------------------|
| NPT4_COMBINE    | Numeric            |average net price for TITLE IV institutions of both public and private.|  Cost                       |
| DEP_STAT_PCT_IND | Numeric           |Percentage of students who are financially independent                |     Cost                        |
| PAR_ED_PCT_1STGEN| Numeric           |Percentage first-generation students                                  |      Sudent Body                |
| MD_FAMINC        | Numeric           |Median family income in real 2015 dollars                             |       Earning                   |
| PCTPELL          | Numeric           |Percentage of undergraduates who receive a Pell Grant                 |        Financial Aid            |
| PCTFLOAN         | Numeric           |Percent of all federal undergraduate students receiving a federal student loan      |  Financial Aid    |
| GRAD_DEBT_MDN    | Numeric           |The median debt for students who have completed                       |   Financial Aid                 |
| WDRAW_DEBT_MDN   | Numeric           |The median debt for sutdents who have not completed                   |Financial Aid |
| MD_EARN_WNE_P6   | Numeric           |Median income six years after entry                                   |     earning   |
| MD_EARN_WNE_P10  | Numeric           |Median income ten years after entry                                   |     earning  |
| RPY_3YR_RT_SUPP  | Numeric           |Repayment rate 3 years after entering repayment                       |Repayment   |
| CDR3             | Numeric           |Loan default rate 3 years afer entering repayment                     |Repayment        |
| variable$qreF    | Factor            |Predominant degree offered                                            |School       |
| variable$ctrF    | Factor            |Control of institution                                                |School     |
| variable$STABBR  | Factor            |State                                                                 |School     |
| APPL_SCH_PCT_GE2 | Numeric           |Percentage of students submitting 2 or more than 2 applications       | Student        |

we will summarize the data set using dplyr functions in the variables summary section.

#4. Objective
#5. Summary of variables
##5.1 Individual variable summary

###5.1.1 School
####5.1.1.1. State
State variable has 59 levels, which include 55 states in the United States and 4 districts outside mainland US. To better understand how many universities each state has, we create a frequency table as shown below, and also a bar chart to visuzalize the top 15 states having the most universities. Some basic summary statistics are also displayed.
```{r, message=FALSE}

state_freq = count(data_clean,"variable$STABBR")[order(-count(data_clean,'variable$STABBR')$freq),]
state_freq
summary(state_freq)
state_freq[1:15,] %>% ggplot(aes(x=factor(variable.STABBR), y=freq)) +
  geom_bar(stat = 'identity')

```
The frequency table counts the number of records belonging to each state, since each records represents a institution, each cell also equals to the number of university in that state.

From frequncy table, we see that California has the most  institutions across the nation, followed by Texas and Ner York. It has about 800 universities! A bar chart is plotted below to show the top 15 states in the frequency table. Then summary of frequency table also shows that the state having the least university only has one, and 50% of state has more than 132 universities in the state.

####5.1.1.2. Predominant Degree
Then we analyze the type of degree universities mainly offer. There are four types of degree taken into accout: Certificate-degree, Associate's degree, Bachelor's degree, and Graduate Degree. A frequency table in a format of percentage and a histogram are used to help understand the distributiton of the Prodominant Degree.

```{r, message=FALSE}
prop.table(table(data_clean$`variable$preF`)) * 100
data_clean %>% ggplot(aes(x=variable$`preF`, fill = 'blue')) + 
  geom_histogram(stat = "count") +
  xlab("Degree") 
```
From above frequency table, we see that about 5% of the population are unclassified in terms of the predomiant degree they offer.  43.2% offer Certificate degree as their predominant degree, 27.3% predominantly offer Bachelor degree and only 4.23% are classified as Graduate degree predominant.

####5.1.1.3. Control of Institution

In addition to State and predominant degree, control of institution is another categorical variable in the dataset. The label of Control is to indicate whether or not a institution is privately owned and whether or not it is a for-profit organization. There are three levels of this variable: Public, Private Nonprofit, and Private-profit. Public school is assumed to be non-profit so it is not further divided into two subgroups. Below we show the percentage summary of each category and a bar char is plotted to facilitate understanding.

@@

About 50% out of all institution are private for-profit university, which is the *most dominant type*. Public and Private Nonprofit are pretty tiex in terms of the number of observation in each category.

###5.1.2 Student Body
####5.1.2.1 Percentage first-generation students
This element identifies the share of federally aided students at the institution who report that they are first generaion on the FAFSA (Free Application for Federal Student Aid). First generation students are defined as sudents in which neither parents completed college.

@@

```
From the output, we can tell that the highest value is 95.25% while the lowest value is 6.87%. From the mean value of 48.08%, we can tell that about for federally aided students in most colleges, around half (50.8%) of the students' parents did not complete the college. The standard deviation value of 12.2% indicated a relative high difference of this element among different colleges.

@@

####5.1.2.2 Number of applications

The variable APPL_SCH_PCT_GE2 is to describe the percentage of students who list two and more than two institutions when he applied for financial aid when submitting application for university. This information is only available to students who submitted their applications for federal aid, and does not cover the whole student body. Nevertheless, this variable still can provides some useful insights about how many universities a student usually apply for.

@@

The summry statistics show that median percentage of student submitting two or more than two application is 41.67% across all univerisities. The institution having the least percentage only has 4% of student applying to more than one university, while the school with the highest percent has as high as 90% of student applying to more than one university. The boxplot compares the the percent of student applying to only one university with that of student applying for more than one. We can clearly see that a larger portion of students only listed one instituiton when submitting their application for financial aid.

###5.1.3 Cost

####5.1.3.1 Average net price for Title IV institutions(both public and private institutions)

We are using the code below to summarize the variable. 
```{R, message=FALSE, warinig = FALSE}

@@

```
From the summary results, we learned that the maximum value of the variable is \$87,414 while the minimum value is about \$-19,360. Considering the elements are collected from Title IV institutions, we reasoned that the negative values may be due to the fact that some intituions's finainal aid supoort can cover the students' all the costs, meanwhile overbalancing the cost amount. There are not much differerences between mean value and median value, which is around \$15,000. We can conclude that the average net price costs for most institutuins are falling into the this range. The standard deviation value is 7604.884 which is relatively large, indicating a great spread in the data.

@@

####5.1.3.2 Percentage of students who are financially independent

Here is the code to generate the summary statistics: 
```{R}

@@

```
From the summary results, we learned that the maximum value of the variable is 98.56% while the minimum value is about 0，73%. From the mean value, we can tell that about in most colleges, around 56% of the students are financially independent. The standatd deviation value of 23.3% indicates a relative high different among different colleges, which can also be concluded from the difference between maximum value and minimum value. Below is the code to generate a visual representation in the histgram. 
```{R}

@@

Then we grouped by the data into the control type(Private For-profit, Public, Private Nonprofit) to take a further analysis.
```{R message=FALSE, warinig = FALSE}

###5.1.4 Financial Aid 
####5.1.4.1 Percentage of undergraduates who receive a Pell Grant

This elemnt, pulled from the IPEDS student Financial Aid Component, shows the share of undergraduate students who received Pell Grants in a given year, which is an important measure of the access an insiitution provides to low-income students. 
We are using the code below to summarize this variable.

@@

@@

```
From the output, we can tell that the range is between 0 to 100%.	 Among the colleges in the sample, half colleges have this rate as high as 52.7% (median value of 52.7%), which means more than half of undergraduare students are able to receive the Pell Grant to cover their costs. The standard deviation value is 22.57%,  indicating a relative high difference of this element among different colleges, which can also be noticed from the largest range (100%). Below is the code to generate a visual representation in the histgram.

@@

####5.1.4.2 Percent of all federal undergraduate students receiving a federal student loan

This elemnt, as reported in the IPEDs Student Finaicial Aid component, show the share of undergraduate students who received federal loans in a given year. It can provide important context to figures related to debt, repayment and non-repayment.  
We are using the code below to summarize this variable.

@@

@@

```
From the output, we can tell that the range is between 0 to 100%.	 Among the colleges in the sample, half colleges have this rate as high as 60.08% , which means around 60% of undergraduare students are able to receive the fedaral student loans. The standard deviation value is 29.26%,  indicating a relative high difference of this element among different colleges, which can also be noticed from the largest range (100%). Below is the code to generate a visual representation in the histgram.

@@

Then we grouped by the data into the college type level to take a further analysis.
```{R message=FALSE, warinig = FALSE}

data_clean %>%
  group_by(variable$STABBR ) %>%
    dplyr::summarise(minPCTFLOAN=min(PCTFLOAN ,na.rm=TRUE),    
    maxPCTFLOAN=max(PCTFLOAN,na.rm=TRUE),medianPCTFLOAN=median(PCTFLOAN,na.rm=TRUE))  %>%
       arrange(medianPCTFLOAN)   %>%
{.}  ->PCTFLOAN_STATE

PCTFLOAN_STATE

#stacked histogram
data_clean %>%
  ggplot(aes(x=PCTFLOAN)) +
  geom_bar(binwidth=0.1,aes(fill=variable$ctrF ))

From the result, the Private institutions have a much larger value than the public institutions. With a much higher costs burden, it makes sence that the low-income sutdents in private institutions are much more likely to seek financial aid than the public institution students.

####5.1.4.3 The median debt for students who have completed

This element is the median loan debt acculated at the institution by all student borrowers of federal lonans who have habe completed the education.We are using the code below to summarize this variable. 
```{R, message=FALSE, warinig = FALSE}

@@

```
From the output, we can tell that the highest value is 48,250	 while the lowest value is 1448. From the median value of 14000, we can tell that among the colleges in the sample, the students who have completed the deducaiton from half of them had the median loan debt of 14000. The standard deviation value is 8020.958,  indicating a relative high difference of this element among different colleges. Below is the code to generate a visual representation in the histgram.

@@

####5.1.4.4 The median debt for students who have not completed

WDRAW_DEBT_MDN describes the median amount of debt fot students who ahve not completed their study. In other words, they are still attending the university.

```{r, message=FALSE}
median_debt_non_null = data_clean$WDRAW_DEBT_MDN[!is.na(data_clean$WDRAW_DEBT_MDN)]
summary(median_debt_non_null)
sd(median_debt_non_null)

@@

```
We use summary statistics, standard deviation and histogram to understand the distribution of this variable.

The summary statistics show that students in 50% of institution have a median debt of more than 5500, which is not very unaffordable. The median debt for attending students can be as much as 27620 for some school and as low as 600 for others, which indicates it may have a large variance. So we also calculate standard deviation to measure the dispersion of the distribution. The standard deviation is 2415.623 about 95% of institution have a median debt ranging from 1167.754  to 10830.246. The histogram shows that the distribution of the median roughly follows a standard distribution with a sharp peak and slightly heavy tail.

###5.1.5 Earning
####5.1.5.1 Median family income in real 2015 dollars
This elemnt is recording the median family income of students, one of the undergraduate student demographics for earnings cohorts.

@@

@@

The median default rate aross all universities is 13%, which means that more than 50% of universities have 13 out of 100 students defaulted their loan!

#6. Relationship between variables
##6.1. NPT4_COMBINE VS State

To analyze the relationship between average net cost for Title IV institutions and State, we grouped the data into the state level to take a further look.
```{R message=FALSE, warinig = FALSE}
